---
title: "Data_manipulation"
output: github_document
date: "2023-03-12"
---

```{r setup, message=FALSE}
library(tidyverse)
```


## load in the FAS Litters Data

```{r}
Litters_df = read_csv("./data/FAS_litters.csv")
Litters_df = janitor::clean_names(Litters_df)
Litters_df
```

## dplyr package Select 

choose some columb but not others

```{r}
select(Litters_df, group, gd0_weight)
select(Litters_df, group, gd0_weight:gd_of_birth)
select(Litters_df, -group)
```

Rename column

```{r}
rename(Litters_df, GROUP = group)
```

select helpers

starts_with only include
```{r}
select(Litters_df, starts_with("gd"))
```

move column to front 

```{r}
select(Litters_df, litter_number, everything())
relocate(Litters_df,litter_number)
```

## filter

remove rows condition based

```{r}
filter(Litters_df, gd0_weight < 22)
```

```{r}
filter(Litters_df, gd0_weight >= 22)
```

```{r}
filter(Litters_df, gd_of_birth == 20)
```
reverse the above

```{r}
filter(Litters_df, !(gd_of_birth == 20))
```

union condition

```{r}
filter(Litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(Litters_df, group == "Mod8")
```

union condition

```{r}
filter(Litters_df, group %in% c("Mod8","Con7"))
```

## mutate

create variable and modify variable 

```{r}
mutate(Litters_df,wt_gain = gd18_weight -gd0_weight, 
       group = str_to_lower(group))
```

## arrange 

arrange by first and by second within first

```{r}
arrange(Litters_df, pups_born_alive)
```

## %>%

pipes: steps to clean data

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```
drop_na(df, droup by what is missing)

USE PIPE
```{r}
litters_df_pipe = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

command shift m : hot key for %>% 