---
title: "File_Import_practice"
output: github_document
---

```{r setup}
library(tidyverse)
```

Set Up

## Read in some data

```{r readin}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


## Different ways to look at the data


```{r}
litters_df
```

view the front and the back of the data use the following functions:

```{r}
head(litters_df)
tail(litters_df)
```

Skim the relevant data summary

```{r}
skimr::skim(litters_df)
```

We can also use view() function in the shell to see the full data

## Options to read_csv

read_csv("", skip = 10, col_names = FALSE)    skip the first 10 obs and rename the col to x1-xn

read_csv("", na = c("", "NA","."))    N/A processing

check our '?read_csv()' for more info

## Other file format

### Read in exacl file

```{r setup_2, message=FALSE}
library(readxl)
```

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

## Options to read_excel

range = "A1:E7"

Sheet = ""

### Read in SAS

```{r}
library(haven)
```

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```
## Comparison with Base R

read.csv vs read_csv 

NEVER USE .csv

## Export data

write_csv(mlb,df, "path")
